Before discussing how FLAM and LIO fit together, we briefly introduce each separately. We first highlight the essential parts of FLAM necessary for this work; more details can be found in \textcite{Arden:2015:FA:2859845.2859998}. Similarly, we highlight the essential parts of LIO, and additional details can be found in \textcite{SRMMlio}.

\subsection{The FLAM principal lattice}
Figure~\ref{fig:flam-syntax} describes the syntax of FLAM principals. The grammar is parametric in a set $\Nameset$ of names representing principals like $\mathsf{Alice}$ and $\mathsf{Bob}$. Given a principal $p$, FLAM can talk about the confidentiality and integrity of $p$ using \emph{basis projections} $\conf{p}$ and $\integ{p}$ respectively. The principal $\conf{p}$ represents the authority to learn anything that $p$ can learn, and $\integ{p}$ represents the authority to modify anything that $p$ can modify. Given principals $p$ and $q$, FLAM can also represent the authority of \emph{both} $p$ and $q$ as $p \conj q$ or the authority of \emph{either} $p$ or $q$ as $p \disj q$. This forms a lattice $(\mathcal{P}, \actsfor)$ with a partial order $\actsfor$ (pronounced \emph{acts for}), that represents trust: if $p \actsfor q$ principal $p$ is allowed to act on behalf of $q$, i.e., $q$ trusts $p$ to act on its behalf. Principals $\bot$ and $\top$ represent the least and most trusted principals respectively, and operations $\conj$ and $\disj$ are the lattice join and meet operations respectively. The principals in $\mathcal{P}$ are given by the grammar in Figure~\ref{fig:flam-syntax}.\footnote{Formally, the principals in the lattice is the equivalence class of $\mathcal{P}$ modulo the relation $\equiv$ where $a \equiv b \iff a \actsfor b$ and $b \actsfor a$.} Besides basis projections $\rightarrow$ and $\leftarrow$, FLAM also defines \emph{ownership projections} $\owner{o}{p}$ representing the same authority as the principal $p$, but where the owner $o$ controls which principals $\owner{o}{p}$ trusts. For example, the delegation $\owner{\mathsf{Acme}}{\mathsf{Bob}} \actsfor \mathsf{Acme}$ implies that $\mathsf{Bob}$ can act for $\mathsf{Acme}$, but it is not the case that $\mathsf{Alice}$ can act for $\owner{\mathsf{Acme}}{\mathsf{Bob}}$, even if $\mathsf{Alice}$ acts for $\mathsf{Bob}$. We will use ownership extensively in the example presented in Section~\ref{subsec:agents}.

\begin{figure}
    \centering
    \begin{tabular}{ll}
    $n \in \Nameset$ \\
    $p ::= \bot \mid \top \mid \name \mid p \conj p \mid p \disj p \mid \conf{p} \mid \integ{p} \mid \owner{p}{p}$
    \end{tabular}
    \caption{Syntax of FLAM}
    \label{fig:flam-syntax}
\end{figure}


\paragraph{An information-flow ordering}
An important distinction between FLAM and other authorization models is that FLAM unifies trust and information-flow into a single concept. That is, in FLAM, principals denote both entities with security concerns and information-flow labels that can be used to restrict the propagation of information in a system. The acts-for ordering describes trust relationships between principals, and is used to define an information-flow ordering that describes the permitted propagation of information. Specifically, FLAM defines the operations
\begin{align}
\label{eq:flam-flowsto-from-actsfor} p \flowsto q &\circeq \conf{q} \conj \integ{p} \actsfor \conf{p} \conj \integ{q}\\
p \join q &\circeq \conf{(p \conj q)} \conj \integ{(p \disj q)} \\
p \meet q &\circeq \conf{(p \disj q)} \conj \integ{(p \conj q)}
\end{align}
That is, $p \flowsto q$ (pronounced $p$ \emph{flows to} $q$) if $q$ acts for the confidentiality of $p$, and $p$ acts for the integrity of $q$. So if $p \flowsto q$, information labeled with principal $p$ can safely be relabeled to principal $q$, as $q$ is at least as confidential, and has no more integrity, than $p$. The \emph{join} of $p$ and $q$, written $p \join q$ is defined as the principal with the authority of both $p$'s and $q$'s confidentiality, and the authority of either $p$'s or $q$'s integrity. The \emph{meet} of $p$ and $q$, written $p \meet q$ is defined dually as the confidentiality of either $p$ or $q$, and the integrity of both $p$ and $q$.
This forms a lattice $(\mathcal{P}, \flowsto)$ with the partial order $\flowsto$, where the bottom element $\botinfoflow \circeq \conf{\bot} \conj \integ{\top}$ represents the least confidential and most trusted principal, and the top element $\topinfoflow \circeq \conf{\top} \conj \integ{\bot}$ represents the most confidential and least trusted principal.\footnote{Once again, the elements is equivalence classes of $\mathcal{P}$ modulo the relation $\equiv$ defined as $a \equiv b \iff a \flowsto b$ and $b \flowsto a$.}

\paragraph{Voice of a principal}
Finally, FLAM defines the \emph{voice} of a principal $p$, denoted $\voice{p}$, as the minimum integrity needed to influence the flow of information labeled $p$. Using the voice operator, FLAM defines a \emph{speaks for} relation \cite{Lampson:1992:ADS:138873.138874, Abadi:2006:ACC:1159803.1159839} between principals as ``principal $p$ speaks for principal $q$ if $p$ acts for the voice of $q$''. Formally $p$ speaks for $q$ if $p \actsfor \voice{q}$. Any principal is equivalent to the conjunction of a confidentiality projection and an integrity projection, i.e., $\forall p.~ \exists q, r.~ p \equiv \conf{q} \conj \integ{r}$. The voice of a principal $p$ with normal form $\conf{q} \conj \integ{r}$ is then defined as $\voice{p} = \integ{q} \conj \integ{r}$ \cite{Arden:2015:FA:2859845.2859998}.

\subsection{Coarse-grained information flow using LIO.}
LIO \cite{SRMMlio} is a Haskell library for dynamic information-flow control. LIO is parametric in the label model and takes a coarse-grained approach to information-flow using a \emph{floating label model}: instead of attaching a label to each value in the program, the \emph{computational context} is protected with a single label called the \emph{current label}. Throughout the execution of the program the current label will ``float up'' the information-flow lattice as more confidential (or less trustworthy) information is brought into the computational context. The current label restricts what data can be modified, ensuring that non-confidential side effects do not depend on confidential information, and dually that trusted effects do not depend on untrusted information.

The type of LIO computations gives rise to a monad \cite{Wadler:1995:MFP:647698.734146} that encapsulates raising the current label. The monadic structure of LIO makes programming with it convenient in Haskell. Specifically, the operation $\return{\expr}$ embeds a pure expression $\expr$ into the LIO computational context, and the operation $\bind{\expr_1}{\expr_2}$ (pronounced \emph{bind}) chains together monadic LIO operations $\expr_1$ and $\expr_2$. Throughout the paper, we use Haskell's \emph{do notation} in examples, which can be desugared into the calculus presented in Section~\ref{sec:calculus}.

In addition to the current label, LIO also provides a \emph{clearance label}, imposing an upper bound on the current label. The clearance label gives LIO a form of access control by restricting which data can be observed and modified within a computational context. For instance, given two labels $\clr$ and $\level_{\mathsf{data}}$ such that $\level_{\mathsf{data}} \nflowsto \clr$, a computation with clearance label $\clr$ cannot access information labeled with label $\level_{\mathsf{data}}$ because that would require the current label to float up to $\level_{\mathsf{data}}$, but $\clr$ is an upper bound on the current label and $\level_{\mathsf{data}} \nflowsto \clr$.

\paragraph{Labeled values}
In addition to protecting every value in the computational context by a single label $\level_{\mathsf{cur}}$, LIO also allows computations to associate explicit labels with particular values. This allows computations to handle data of different labels in the same context. The type of labeled values of type \texttt{a} is written in Haskell as \mintinline{haskell}{Labeled l a}, where \texttt{l} is the type of labels.

Labeled values are typically used to incorporate sensitive information such as usernames and passwords into the computational context. LIO provides three operations for working with labeled values:
\begin{minted}{haskell}
label :: Label l => a -> l -> LIO l (Labeled l a)
unlabel :: Label l => Labeled l a -> LIO l a
labelOf :: Label l => Labeled l a -> l
\end{minted}
Here, \mintinline{haskell}{Label l} is a typeclass constraint specifying that the type \texttt{l} must be an instance of the \mintinline{haskell}{Label} typeclass, meaning that \texttt{l} must have operations $\join$, $\meet$ and $\flowsto$. The operation \mintinline{haskell}{label} takes an expression $\expr$ and a label $\level$, and labels $\expr$ with the label $\level$. When the value of the labeled value is needed, the operation \mintinline{haskell}{unlabel} must be invoked, which gets back the value and raises the current label $\level_{\mathsf{cur}}$ to $\level_{\mathsf{cur}} \join \level$, while checking that the new current label flows to the clearance label $\level_{\mathsf{clr}}$ of the computation.

Finally, \mintinline{haskell}{labelOf} extracts the label of a labeled value. This operation does not return a value in the LIO monad, and thus no information-flow checks are performed when invoking \mintinline{haskell}{labelOf}. In other words, the label is protected only by the current label \cite{SRMMlio}. This fact is important when we define trust checking using strategies in Section~\ref{subsec:deriving-trust}.

\paragraph{Preventing label creep}
As the program executes and confidential or untrusted information enters the computational context through \mintinline{haskell}{unlabel} operations, the current label continues to creep upwards, restricting the possible side effects. To avoid unnecessary label creep LIO introduces the following operation:
\begin{minted}{haskell}
toLabeled :: Label l => l -> LIO l a ->
               LIO l (Labeled l a)
\end{minted}
Evaluating \mintinline{haskell}{toLabeled l e} when the current label is $\level_{\mathsf{cur}}$ will evaluate \texttt{e} and then reset the current label to $\level_{\mathsf{cur}}$. To remain secure, the result is labeled with the label \texttt{l}. Furthermore, LIO checks that the evaluation of \texttt{e} never raises the current label above \texttt{l}.